\mytitle

### 2023-06-06 expanded-paper.tex
Preparing to upload to Computer Journal

\documentclass{comjnl}
\usepackage{ragged2e}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\input paper-seb-macros.tex
\input generated/page-lengths.tex

\makeatletter
      \input paper-seb-supplementary-material.aux 
\makeatother

>>>>

% 
% This file has been expanded from LaTeX file (paper.tex)
%   to make this straight forward LaTeX file (expanded-paper.tex)
%     that depends on nothing (it inputs no files apart from the document class files).
%       Each required file's text was input in situ - search for % expand
% 
% (Otherwise the file paper.tex would use BibTeX, a macro file, tables of contents, its own and the other files's .aux files, and includes various files in generated/*, etc.)
%

\makeatletter
\relax 
\providecommand\hyper@newdestlabel[2]{}
\providecommand\HyperFirstAtBeginDocument{\AtBeginDocument}
\HyperFirstAtBeginDocument{\ifx\hyper@anchor\@undefined
\global\let\oldnewlabel\newlabel
\gdef\newlabel#1#2{\newlabelxx{#1}#2}
\gdef\newlabelxx#1#2#3#4#5#6{\oldnewlabel{#1}{{#2}{#3}}}
\AtEndDocument{\ifx\hyper@anchor\@undefined
\let\newlabel\oldnewlabel
\fi}
\fi}
\global\let\hyper@last\relax 
\gdef\HyperFirstAtBeginDocument#1{#1}
\providecommand\HyField@AuxAddToFields[1]{}
\providecommand\HyField@AuxAddToCoFields[2]{}
\citation{a=b}
\citation{write-for-life}
\citation{abelson}
\citation{nature-chatgpt}
\newlabel{FirstPage}{{}{1}{}{Doc-Start}{}}
\@writefile{toc}{\contentsline {section}{\numberline {1}Introduction}{1}{section.1}\protected@file@percent }
\citation{sommerville,knight}
\citation{computational-thinking-wing,computational-thinking}
\citation{ICmodel,avianFluModel,originalICmodel}
\citation{ICmodel,avianFluModel,originalICmodel}
\citation{reproducibility-crisis,rescience}
\@writefile{toc}{\contentsline {section}{\numberline {2}Background}{2}{section.2}\protected@file@percent }
\@writefile{lot}{\contentsline {table}{\numberline {1}{\ignorespaces Overview of peer-reviewed paper sample, broken down further in Table \ref {table-summary}. Survey methodology and data is provided in the Supplemental Material. (The survey does not include the motivating papers \cite  {ICmodel,avianFluModel,originalICmodel}, none of which provide code or code summaries; see section \ref {section-pandemic-modeling}.)}}{3}{table.1}\protected@file@percent }
\newlabel{table-overview}{{1}{3}{Overview of peer-reviewed paper sample, broken down further in Table \ref {table-summary}. Survey methodology and data is provided in the \supplement . (The survey does not include the motivating papers \cite {ICmodel,avianFluModel,originalICmodel}, none of which provide code or code summaries; see section \ref {section-pandemic-modeling}.)}{table.1}{}}
\@writefile{lot}{\contentsline {table}{\numberline {2}{\ignorespaces Breakdown of pilot survey of peer-reviewed science papers relying on code. }}{3}{table.2}\protected@file@percent }
\newlabel{table-summary}{{2}{3}{Breakdown of pilot survey of peer-reviewed science papers relying on code}{table.2}{}}
\@writefile{lot}{\contentsline {table}{\numberline {3}{\ignorespaces The TOP committee's recommended levels for journal article code transparency. Level 0 is provided for a comparison that does not meet any TOP requirements. Concerns about the interpretation of ``reproduced independently,'' as required at level 4, are raised in section \ref {reproducibility-concerns}.}}{3}{table.3}\protected@file@percent }
\newlabel{TOPtable}{{3}{3}{The TOP committee's recommended levels for journal article code transparency. Level 0 is provided for a comparison that does not meet any TOP requirements. Concerns about the interpretation of ``reproduced independently,'' as required at level 4, are raised in section \ref {reproducibility-concerns}}{table.3}{}}
\citation{temperature}
\citation{form}
\citation{acm-algorithms}
\citation{hoare-correctness}
\citation{jupyter-study}
\citation{trisovic}
\citation{jupyter-study}
\citation{trisovic}
\citation{trisovic}
\@writefile{toc}{\contentsline {subsection}{\numberline {2.a}Code quality concerns}{4}{subsection.2.1}\protected@file@percent }
\newlabel{related-work}{{2.a}{4}{Code quality concerns}{subsection.2.1}{}}
\citation{machine-learning-reproducibility,no-raw-data,data-access,reproducible-manifesto,reproducibility-crisis}
\citation{joss}
\citation{rescience}
\citation{TOP}
\citation{TOP-wiki}
\citation{jupyter-study}
\citation{MMR-fraud}
\citation{misconduct-issues}
\citation{machine-learning-reproducibility}
\citation{elns}
\@writefile{toc}{\contentsline {subsection}{\numberline {2.b}Computable papers}{5}{subsection.2.2}\protected@file@percent }
\newlabel{computable-papers}{{2.b}{5}{Computable papers}{subsection.2.2}{}}
\citation{ICmodel}
\citation{perkel-review}
\citation{maneage}
\citation{maneage}
\citation{LP,LP-book}
\citation{relit}
\citation{notebook-patent}
\citation{jupyter-history}
\citation{knitr}
\citation{spec-led-design}
\citation{RAPstandard}
\@writefile{toc}{\contentsline {subsection}{\numberline {2.c}RAP: Reproducible Analytical Pipelines}{7}{subsection.2.3}\protected@file@percent }
\newlabel{RAP-section}{{2.c}{7}{RAP: Reproducible Analytical Pipelines}{subsection.2.3}{}}
\citation{rap,goldacre,turing-way}
\@writefile{lot}{\contentsline {table}{\numberline {4}{\ignorespaces A simplified schematic of the publication pipeline. For clarity, the pipeline has been linearized; in general, there will be repetitive cyclic iteration and refinement. The RAP and RAP{\tt  +}\ approaches encode the normally manual steps in the pipeline workflow so that they can be run automatically, and hence reproduce the results that underpin the final paper. The encoded RAP{\tt  +}\ algorithms can be shared with other scientists, scrutinized, simplified and optimized, and themselves turned into publishable objects --- they are scientific instruments, just like thermometers or DNA sequencers. Additional schematics are provided in section \ref {supplement-RAP-diagrams} in the Supplemental Material.}}{8}{table.4}\protected@file@percent }
\newlabel{fig-pipeline}{{4}{8}{A simplified schematic of the publication pipeline. For clarity, the pipeline has been linearized; in general, there will be repetitive cyclic iteration and refinement. The RAP and \RAPstar \ approaches encode the normally manual steps in the pipeline workflow so that they can be run automatically, and hence reproduce the results that underpin the final paper. The encoded \RAPstar \ algorithms can be shared with other scientists, scrutinized, simplified and optimized, and themselves turned into publishable objects --- they are scientific instruments, just like thermometers or DNA sequencers. Additional schematics are provided in section \ref {supplement-RAP-diagrams} in the \supplement }{table.4}{}}
\@writefile{lot}{\contentsline {table}{\numberline {5}{\ignorespaces A minimum standard of RAP, based on the UK Statistics Authority summary \cite  {RAPstandard}.}}{8}{table.5}\protected@file@percent }
\newlabel{figure-RAP-summary}{{5}{8}{A minimum standard of RAP, based on the UK Statistics Authority summary \cite {RAPstandard}}{table.5}{}}
\citation{MMR-fraud}
\citation{a=b}
\citation{Courtes}
\@writefile{toc}{\contentsline {subsubsection}{\numberline {2.c.1}Potential code}{9}{subsubsection.2.3.1}\protected@file@percent }
\@writefile{toc}{\contentsline {subsubsection}{\numberline {2.c.2}RAP as research}{9}{subsubsection.2.3.2}\protected@file@percent }
\citation{Courtes}
\citation{Courtes}
\citation{apa-statistics,Mariusz-reporting,science-statistics-recommendations}
\citation{example-stats}
\citation{science-statistics-recommendations}
\citation{Spiegelhalter}
\citation{paul-hci,pnas-stats}
\citation{paul-hci,pnas-stats}
\citation{Ben,se-bias}
\@writefile{toc}{\contentsline {section}{\numberline {3}The statistics/code analogy}{10}{section.3}\protected@file@percent }
\newlabel{statistics-computation-analogy}{{3}{10}{The statistics/code analogy}{section.3}{}}
\citation{sas9.2}
\citation{whitty}
\newlabel{section-vicious-cycles}{{3}{11}{The statistics/code analogy}{section.3}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {3.a}The siren call of over-fitting code}{11}{subsection.3.1}\protected@file@percent }
\newlabel{over-fit}{{3.a}{11}{The siren call of over-fitting code}{subsection.3.1}{}}
\citation{over-fit}
\citation{over-fit}
\citation{may-simple}
\citation{dyson}
\newlabel{name-example-footnote}{{5}{12}{}{Hfootnote.5}{}}
\citation{assurance-case}
\@writefile{lof}{\contentsline {figure}{\numberline {1}{\ignorespaces Much computational science is concerned with finding plausible multi-dimensional models that fit models to data with the aim of extrapolating or predicting new results from them. Shown here is notional sample of experimental 2D data (the dots), a linear least squares regression, and an exact polynomial model. The over-fitted polynomial model fits the sample \emph  {exactly\/}, but since the experimental data is presumably subject to random error (indicated by the confidence interval, itself estimated) the linear model would generally be considered a better description of the experimental data.}}{13}{figure.1}\protected@file@percent }
\newlabel{fig-overfit}{{1}{13}{Much computational science is concerned with finding plausible multi-dimensional models that fit models to data with the aim of extrapolating or predicting new results from them. Shown here is notional sample of experimental 2D data (the dots), a linear least squares regression, and an exact polynomial model. The over-fitted polynomial model fits the sample \emph {exactly\/}, but since the experimental data is presumably subject to random error (indicated by the confidence interval, itself estimated) the linear model would generally be considered a better description of the experimental data}{figure.1}{}}
\@writefile{toc}{\contentsline {section}{\numberline {4}The conventional role of code}{13}{section.4}\protected@file@percent }
\citation{assessing-quality}
\citation{fixit}
\citation{Hammurabi}
\@writefile{toc}{\contentsline {subsection}{\numberline {4.a}The deceptive simplicity of code}{14}{subsection.4.1}\protected@file@percent }
\newlabel{deceptive-simplicity-of-code}{{4.a}{14}{The deceptive simplicity of code}{subsection.4.1}{}}
\citation{cite-code,TOP}
\citation{prisma}
\citation{engima}
\citation{critiques,diagnosis-reviews}
\@writefile{toc}{\contentsline {subsection}{\numberline {4.b}The low status of coding}{15}{subsection.4.2}\protected@file@percent }
\newlabel{lowCodeStatus}{{4.b}{15}{The low status of coding}{subsection.4.2}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {4.c}The critical role of code is often ignored}{15}{subsection.4.3}\protected@file@percent }
\newlabel{central-role-of-code}{{4.c}{15}{The critical role of code is often ignored}{subsection.4.3}{}}
\newlabel{PRISMA-statement}{{4.c}{15}{The critical role of code is often ignored}{subsection.4.3}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {4.d}Bugs, code and programming}{15}{subsection.4.4}\protected@file@percent }
\newlabel{knowledge}{{4.d}{15}{Bugs, code and programming}{subsection.4.4}{}}
\citation{knuth}
\citation{essence-of-software,fixit}
\citation{popper-conjectures-refutations}
\citation{preventable-book}
\@writefile{toc}{\contentsline {subsection}{\numberline {4.e}Long-term problems of unreliable code}{16}{subsection.4.5}\protected@file@percent }
\citation{preventable-book}
\citation{science-review}
\citation{diagnosis-reviews}
\citation{tripod}
\citation{diagnosis-reviews}
\citation{diagnosis-reviews}
\citation{nature-summary,ICmodel}
\citation{avianFluModel,originalICmodel}
\citation{ICmodel,avianFluModel,originalICmodel}
\citation{tweet}
\citation{ferguson-interview}
\citation{basic-reproducibilty}
\citation{ferguson-resign}
\@writefile{toc}{\contentsline {section}{\numberline {5}State of the art}{17}{section.5}\protected@file@percent }
\@writefile{toc}{\contentsline {subsection}{\numberline {5.a}Case study: Pandemic modeling}{17}{subsection.5.1}\protected@file@percent }
\newlabel{section-pandemic-modeling}{{5.a}{17}{Case study: Pandemic modeling}{subsection.5.1}{}}
\citation{tweet}
\citation{forsyth,Normal-Distributions}
\citation{misra}
\citation{spark}
\citation{haskell}
\citation{thumbs-up}
\citation{nature-summary,ICmodel}
\citation{codecheck,thumbs-up}
\citation{codecheck}
\citation{ICmodel}
\citation{codecheck-certificate}
\citation{codecheck}
\citation{codecheck-certificate}
\newlabel{discuss-covid-sim}{{5.a}{18}{Case study: Pandemic modeling}{subsection.5.1}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {5.b}Concerns with reproducibility}{18}{subsection.5.2}\protected@file@percent }
\newlabel{reproducibility-concerns}{{5.b}{18}{Concerns with reproducibility}{subsection.5.2}{}}
\citation{codecheck,thumbs-up}
\citation{codecheck-certificate}
\citation{NVP}
\citation{nvp-ferguson}
\citation{essence-of-software}
\citation{notebooks,popper-conjectures-refutations,tripod,prisma,nih-policy,nih-nature}
\citation{refactoring}
\citation{bad-code}
\citation{bad-code}
\citation{pseudo}
\citation{chinese}
\citation{relit}
\citation{flu-model}
\citation{nature-review}
\citation{nature-review}
\citation{psychological-modeling}
\citation{psychological-modeling}
\citation{actr}
\citation{psychological-modeling}
\citation{Ben,nature-review,se-bias}
\citation{ourReview}
\citation{hamming}
\citation{NVP}
\@writefile{toc}{\contentsline {subsection}{\numberline {5.c}Beyond pandemic modeling}{20}{subsection.5.3}\protected@file@percent }
\newlabel{section-science-beyond-pandemic-modeling}{{5.c}{20}{Beyond pandemic modeling}{subsection.5.3}{}}
\citation{unfinished}
\citation{unfinished,lancet-unfinished}
\citation{hoare}
\citation{machine-learning-reproducibility}
\citation{relit}
\citation{preclinical-reproducibility}
\citation{fair,fair-principles}
\citation{nih-policy}
\citation{nih-nature}
\newlabel{section-discussion}{{5.c}{21}{Beyond pandemic modeling}{subsection.5.3}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {5.d}The lop-sided emphasis on data}{21}{subsection.5.4}\protected@file@percent }
\newlabel{critique-fair}{{5.d}{21}{The lop-sided emphasis on data}{subsection.5.4}{}}
\citation{notebooks}
\citation{cbc}
\citation{ethics-code}
\@writefile{toc}{\contentsline {section}{\numberline {6}Rethinking science that uses code}{22}{section.6}\protected@file@percent }
\newlabel{summary}{{6}{22}{Rethinking science that uses code}{section.6}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {6.a}Software Engineering Boards}{22}{subsection.6.1}\protected@file@percent }
\citation{acm-artifacts}
\citation{redmill,iec61508}
\citation{hoare-unreasonable}
\citation{basic-reproducibilty,open-source}
\citation{ABCs-SE}
\citation{mrb}
\citation{mrb}
\citation{relit,Executable-Paper}
\citation{computational-thinking-wing,computational-thinking}
\@writefile{toc}{\contentsline {subsection}{\numberline {6.b}Extending RAP to RAP{\tt  +}}{23}{subsection.6.2}\protected@file@percent }
\newlabel{RAPstar-section}{{6.b}{23}{Extending RAP to \RAPstar }{subsection.6.2}{}}
\citation{RMarkdown}
\citation{lepton}
\@writefile{toc}{\contentsline {subsection}{\numberline {6.c}The paper as a scientific laboratory}{24}{subsection.6.3}\protected@file@percent }
\newlabel{paper-as-lab}{{6.c}{24}{The paper as a scientific laboratory}{subsection.6.3}{}}
\citation{landauer,thimbleby-landauer}
\@writefile{toc}{\contentsline {subsection}{\numberline {6.d}Action must be interdisciplinary}{25}{subsection.6.4}\protected@file@percent }
\@writefile{toc}{\contentsline {subsection}{\numberline {6.e}Methodological statements}{25}{subsection.6.5}\protected@file@percent }
\citation{no-raw-data}
\citation{data-access}
\citation{assurance-case}
\citation{debt}
\newlabel{mentionPLOSandIEEE}{{6.e}{26}{Methodological statements}{subsection.6.5}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {6.f}Training to reduce technical debt}{26}{subsection.6.6}\protected@file@percent }
\newlabel{technical-debt}{{6.f}{26}{Training to reduce technical debt}{subsection.6.6}{}}
\citation{fixit}
\citation{jpm}
\citation{jpm}
\@writefile{toc}{\contentsline {subsection}{\numberline {6.g}Benefits beyond science}{27}{subsection.6.7}\protected@file@percent }
\newlabel{benefits-beyond-science}{{6.g}{27}{Benefits beyond science}{subsection.6.7}{}}
\citation{jpm-report}
\citation{jpm}
\citation{hoare-unreasonable}
\citation{excel-fiasco}
\citation{science-delays}
\@writefile{toc}{\contentsline {subsection}{\numberline {6.h}Approaches to further work}{28}{subsection.6.8}\protected@file@percent }
\@writefile{toc}{\contentsline {section}{\numberline {7}Conclusions}{28}{section.7}\protected@file@percent }
\citation{tweet,ferguson-interview}
\citation{whitty}
\citation{fixit}
\citation{reproducibility,relit,popper-conjectures-refutations}
\citation{parliamentary-evidence,my-parliamentary-evidence}
\bibstyle{revisedCompJ}
\bibdata{paper.bib}
\bibcite{a=b}{1}
\bibcite{write-for-life}{2}
\bibcite{abelson}{3}
\bibcite{nature-chatgpt}{4}
\bibcite{sommerville}{5}
\bibcite{knight}{6}
\bibcite{computational-thinking-wing}{7}
\bibcite{computational-thinking}{8}
\bibcite{ICmodel}{9}
\bibcite{avianFluModel}{10}
\bibcite{originalICmodel}{11}
\bibcite{reproducibility-crisis}{12}
\bibcite{rescience}{13}
\bibcite{temperature}{14}
\bibcite{form}{15}
\bibcite{acm-algorithms}{16}
\bibcite{hoare-correctness}{17}
\bibcite{jupyter-study}{18}
\bibcite{trisovic}{19}
\bibcite{machine-learning-reproducibility}{20}
\bibcite{no-raw-data}{21}
\bibcite{data-access}{22}
\bibcite{reproducible-manifesto}{23}
\bibcite{joss}{24}
\bibcite{TOP}{25}
\bibcite{TOP-wiki}{26}
\bibcite{MMR-fraud}{27}
\bibcite{misconduct-issues}{28}
\bibcite{elns}{29}
\bibcite{perkel-review}{30}
\bibcite{maneage}{31}
\bibcite{LP}{32}
\bibcite{LP-book}{33}
\bibcite{relit}{34}
\bibcite{notebook-patent}{35}
\bibcite{jupyter-history}{36}
\bibcite{knitr}{37}
\bibcite{spec-led-design}{38}
\bibcite{RAPstandard}{39}
\bibcite{rap}{40}
\bibcite{goldacre}{41}
\bibcite{turing-way}{42}
\bibcite{Courtes}{43}
\bibcite{apa-statistics}{44}
\bibcite{Mariusz-reporting}{45}
\bibcite{science-statistics-recommendations}{46}
\bibcite{example-stats}{47}
\bibcite{Spiegelhalter}{48}
\bibcite{paul-hci}{49}
\bibcite{pnas-stats}{50}
\bibcite{Ben}{51}
\bibcite{se-bias}{52}
\bibcite{sas9.2}{53}
\bibcite{whitty}{54}
\bibcite{over-fit}{55}
\bibcite{may-simple}{56}
\bibcite{dyson}{57}
\bibcite{assurance-case}{58}
\bibcite{assessing-quality}{59}
\citation{assessing-quality-conference}
\bibcite{fixit}{60}
\bibcite{Hammurabi}{61}
\bibcite{cite-code}{62}
\bibcite{prisma}{63}
\bibcite{engima}{64}
\bibcite{critiques}{65}
\bibcite{diagnosis-reviews}{66}
\bibcite{knuth}{67}
\bibcite{essence-of-software}{68}
\bibcite{popper-conjectures-refutations}{69}
\bibcite{preventable-book}{70}
\bibcite{science-review}{71}
\bibcite{tripod}{72}
\bibcite{nature-summary}{73}
\bibcite{tweet}{74}
\bibcite{ferguson-interview}{75}
\bibcite{basic-reproducibilty}{76}
\bibcite{ferguson-resign}{77}
\bibcite{forsyth}{78}
\bibcite{Normal-Distributions}{79}
\bibcite{misra}{80}
\bibcite{spark}{81}
\bibcite{haskell}{82}
\bibcite{thumbs-up}{83}
\bibcite{codecheck}{84}
\bibcite{codecheck-certificate}{85}
\bibcite{NVP}{86}
\bibcite{nvp-ferguson}{87}
\bibcite{notebooks}{88}
\bibcite{nih-policy}{89}
\bibcite{nih-nature}{90}
\bibcite{refactoring}{91}
\bibcite{bad-code}{92}
\bibcite{pseudo}{93}
\bibcite{chinese}{94}
\bibcite{flu-model}{95}
\bibcite{nature-review}{96}
\bibcite{psychological-modeling}{97}
\bibcite{actr}{98}
\bibcite{ourReview}{99}
\bibcite{hamming}{100}
\bibcite{unfinished}{101}
\bibcite{lancet-unfinished}{102}
\bibcite{hoare}{103}
\bibcite{preclinical-reproducibility}{104}
\bibcite{fair}{105}
\bibcite{fair-principles}{106}
\bibcite{cbc}{107}
\bibcite{ethics-code}{108}
\bibcite{acm-artifacts}{109}
\bibcite{redmill}{110}
\bibcite{iec61508}{111}
\bibcite{hoare-unreasonable}{112}
\bibcite{open-source}{113}
\bibcite{ABCs-SE}{114}
\bibcite{mrb}{115}
\bibcite{Executable-Paper}{116}
\bibcite{RMarkdown}{117}
\bibcite{lepton}{118}
\bibcite{landauer}{119}
\bibcite{thimbleby-landauer}{120}
\bibcite{debt}{121}
\bibcite{jpm}{122}
\bibcite{jpm-report}{123}
\bibcite{excel-fiasco}{124}
\bibcite{science-delays}{125}
\bibcite{reproducibility}{126}
\bibcite{parliamentary-evidence}{127}
\bibcite{my-parliamentary-evidence}{128}
\citation{parliamentary-evidence}
\bibcite{assessing-quality-conference}{129}
\def\continueSectionNumbers{7}
\def\continueTableNumbers{5}
\def\continueFigureNumbers{1}
\def\continueFootnoteNumbers{13}
\def\continuePageNumbers{35}
\newlabel{LastPage}{{}{35}}
\gdef \@abspage@last{35}

\makeatother
\documentclass{comjnl}
\usepackage{ragged2e}
\usepackage{amsmath}
\usepackage{amssymb}

@@

% expand macros.tex
% due to finding out that PDF doesn't work everywhere, here's where we define the image type we're using for imaged created by the Mathematica notebooks
\def\imageFileSuffix{jpg}

\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{longtable}
\usepackage{cite}

\renewcommand\thesubsection{\thesection.\alph{subsection}}
\renewcommand\thesubsubsection{\thesection.\alph{subsection}.\arabic{subsubsection}}

\def\httpURL#1{\href{http://#1}{\textcolor{blue}{#1}}}
\def\LONGhttpURL#1#2{\href{http://#1}{\textcolor{blue}{#2}}}

\makeatletter
    \def\do@url@hyp{\do/}
    \usepackage{moreverb}
\makeatother

\DeclareUrlCommand\doi{\def\UrlLeft{{\textrm{DOI}\ } }}
\DeclareUrlCommand\url{\def\UrlLeft{{\textrm{URL}\ } }}
\DeclareUrlCommand\midurl{\def\UrlLeft{}\allowbreak }

\renewcommand{\thefootnote}{\arabic{footnote}}

\def\mytitle{Improving science that uses code}
\def\supplement{Supplemental Material}

\def\citeeg#1{[e.g., \citenum{#1}]}

% calculate rounded percentages
\newcount \t
\newcount \tu
\def\pc#1#2{\t = #1%
\multiply \t by 100000%
\divide \t by #2% ?? 0 times percentage
\tu = \t
\divide \t by 1000% percentage, but no idea of remainder
\multiply \t by 1000% 1000 times percentage, with 00 as bottom digits
\advance \tu by -\t% bottom two digits
\divide \t by 1000\relax%
% round to nearest integer
% if ending in > .5 round up
% if ending in exactly .5 round towards nearest even number
\ifnum \tu > 500 % round up
	\advance \t by 1 
\else
	\ifnum \tu = 500 
		\ifodd \t % it's odd, so round up to even number 
			\advance \t by 1
		\else % leave rounded to even number 	
		\fi
	\fi
\fi
\the\t\%}

% code to make digits into blocks of 3 (as in 1000 -> 1 000)
\def\xBlockSep{\textcolor{red}{\hskip 0.3em :}\hskip 0.3em} % International Standards are to use a small space. UK typically uses a comma
% NOTE currently, the few decimal numbers (eg 0.000 5) use hskip explicitly, since defining \xBlockSep as a comma would definitely be wrong for decimals
\newcount\xdigitBlockn
\newcount\xdigitBlocktmp
\def\xdigitBlock#1{%
        \xdigitBlockn=#1\relax
        \ifnum \xdigitBlockn < 0
                $-$%
                \xdigitBlockn = -\xdigitBlockn
        \fi
        \let \xdigitBlockpptemp \empty
        \loop \ifnum \xdigitBlockn > 999
                \xdigitBlocktmp = \xdigitBlockn
                \divide \xdigitBlockn by 1000
                \xdigitBlocktmp = \numexpr \xdigitBlocktmp - 1000*\xdigitBlockn \relax
                \edef \xdigitBlockpptemp{\xBlockSep\ifnum \xdigitBlocktmp < 100  0\ifnum \xdigitBlocktmp < 10 0\fi\fi \the\xdigitBlocktmp \xdigitBlockpptemp}%
        \repeat
        \the\xdigitBlockn
        \xdigitBlockpptemp
}

\def\digitBlock#1{\hbox{\xdigitBlock{#1}}}

% simple macro to use a number register to pluralize (add an s) to words....
% \plural{\t}{fred} => \the\t\ freds if plural, or => one fred if singular
\def\plural#1#2{\ifnum #1=1 
	one #2%
\else
	\the#1\ #2s%
\fi}

% initialiseBibliography <title> <starting number> <introductory text>
\newcount \bibciten \bibciten=0
\def\initialiseBibliography#1#2#3{% title, starting citation number, an initial paragraph
	\global\def\refname{\arabic{section} #1}
    \global\def\unnumberedrefname{#1}
    \global\bibciten=#2
    \global\def\startBibliography{#3}
}

\def\bibskip{\vskip 1ex} % the gap between bib items generated by data.js

% read in any constants defined from the JSON or Git data
% expand generated/constants.tex
% Generated by programs/data.js
\def\journalBreakdown{\emph{Lancet Digital Health\/} ($N=6$), \emph{Nature Digital Medicine\/} ($N=12$) and \emph{Royal Society Open Science\/} ($N=14$)}
\def\tabularJournalBreakdown{&\hbox to 3em {\hfill 6}\hskip 1em \emph{Lancet Digital Health}\\&\hbox to 3em {\hfill 12}\hskip 1em \emph{Nature Digital Medicine}\\&\hbox to 3em {\hfill 14}\hskip 1em \emph{Royal Society Open Science}\\}
\global\newcount \dataN \global\dataN=32
\global\newcount \countAuthors \global\countAuthors=264
\global\newcount \countHasBreach \global\countHasBreach=11
\global\newcount \countHasPolicy \global\countHasPolicy=26
\global\newcount \countUsesVersionControlRepository \global\countUsesVersionControlRepository=10
\global\newcount \counthasDataRepository \global\counthasDataRepository=9
\global\newcount \countNoCodeInRepo \global\countNoCodeInRepo=1
\global\newcount \numberOfJournals \global\numberOfJournals=3
\global\newcount \countCodetested \global\countCodetested=0
\global\newcount \hasDevelopedRigorously \global\hasDevelopedRigorously=0
\global\newcount \countFields \global\countFields=36
\global\newcount \countCommentedOut \global\countCommentedOut=7
\global\newcount \countNotCommentedOut \global\countNotCommentedOut=2
\global\newcount \countMissingCode \global\countMissingCode=1
\global\def \averagePagesPerAuthor{1.2916666666666667}
% end expanding generated/constants.tex
% expand generated/metadata.tex
% shell script (make-metadata.sh) generated LaTeX variable values counting errors detected by data.js
\newcount \JSONerrorCount
\JSONerrorCount = 30
\newcount \dataVariableCount
\dataVariableCount = 63
% end expanding generated/metadata.tex
% expand generated/covidsimSummary.tex
% data calculated from the Ferguson covid-sim repo
\def \covidsimkLOC {25}
\def \covidsimfiles {229}
\def \covidsimdata{734 Mb}
% end expanding generated/covidsimSummary.tex
% expand generated/clone-date.tex
% date generated by running generated/allGitRepos.sh downloading Git repos
\def\clonedate {\ignorespaces 25 May 2023}
\def\cloneyear {2023}
\def\clonemonth {05}
% end expanding generated/clone-date.tex

\newcount \pubdelayinmonths
\pubdelayinmonths=\cloneyear
\advance \pubdelayinmonths by -2019 % submitted in 2019
\multiply \pubdelayinmonths by 12
\advance \pubdelayinmonths by -6 % submitted in June
\advance \pubdelayinmonths by \clonemonth

\def\monthname#1{\ifcase#1 \or 
January\or 
February\or 
March\or 
April\or 
May\or 
June\or 
July\or 
August\or 
September\or 
October\or 
November\or 
December\fi}

\def\numbertoname#1{\ifcase #1 zero\or one\or two\or three\or four\or five\or six\or seven\or eight\or nine\else #1\fi}
\def\Numbertoname#1{\ifcase #1 Zero\or One\or Two\or Three\or Four\or Five\or Six\or Seven\or Eight\or Nine\else #1\fi}

\def\clonewhen{\monthname{\clonemonth} \cloneyear}

\def\RAPstar{RAP{\tt +}}
\def\RAPstars{RAPs{\tt +}}
\def\RAPstarp#1{RAP{\tt +}#1}
\def\italicRAPstar{RAP{\tt +}}
% end expanding macros.tex
% expand generated/page-lengths.tex
% Generated by programs/data.js
\expandafter\def\csname pagelength-MetricSelectionFramework\endcsname{17}
\expandafter\def\csname pagelength-PENet\endcsname{9}
\expandafter\def\csname pagelength-PostoperativeOutcomes.RiskNet\endcsname{10}
\expandafter\def\csname pagelength-philter-ucsf\endcsname{8}
\expandafter\def\csname pagelength-AI-CDSS-Cardiovascular-Silo\endcsname{6}
\expandafter\def\csname pagelength-SiameseChange\endcsname{9}
\expandafter\def\csname pagelength-lactModel\endcsname{13}
\expandafter\def\csname pagelength-LRM\endcsname{22}
\expandafter\def\csname pagelength-manifold-ga\endcsname{7}
\expandafter\def\csname pagelength-blast-ct\endcsname{8}
\newcount \gitPages \gitPages=109
\newcount \totalPages \totalPages=341
% end expanding generated/page-lengths.tex

\makeatletter
      % expand appendix.aux 
\relax 
\providecommand\hyper@newdestlabel[2]{}
\providecommand\HyperFirstAtBeginDocument{\AtBeginDocument}
\HyperFirstAtBeginDocument{\ifx\hyper@anchor\@undefined
\global\let\oldnewlabel\newlabel
\gdef\newlabel#1#2{\newlabelxx{#1}#2}
\gdef\newlabelxx#1#2#3#4#5#6{\oldnewlabel{#1}{{#2}{#3}}}
\AtEndDocument{\ifx\hyper@anchor\@undefined
\let\newlabel\oldnewlabel
\fi}
\fi}
\global\let\hyper@last\relax 
\gdef\HyperFirstAtBeginDocument#1{#1}
\providecommand\HyField@AuxAddToFields[1]{}
\providecommand\HyField@AuxAddToCoFields[2]{}
\citation{vancouver}
\citation{hamming}
\citation{fixit}
\citation{cosmos}
\@writefile{toc}{\contentsline {section}{\numberline {8}Further issues for Software Engineering Boards (SEBs)}{38}{section.8}\protected@file@percent }
\@writefile{toc}{\contentsline {subsection}{\numberline {8.a}Relationships of SEBs to Ethics Boards}{38}{subsection.8.1}\protected@file@percent }
\citation{ethics-paper}
\citation{ai-ethics}
\citation{science-lancet1,science-lancet2,lancet-learning}
\citation{jvs1,jvs2,jvs3}
\citation{sommerville,knight}
\citation{cosmos}
\citation{swebok}
\citation{turing-way}
\citation{wpf}
\citation{humphrey}
\citation{managers}
\@writefile{toc}{\contentsline {subsection}{\numberline {8.b}SEBs are necessary but not sufficient}{39}{subsection.8.2}\protected@file@percent }
\@writefile{toc}{\contentsline {section}{\numberline {9}Software engineering best practice}{39}{section.9}\protected@file@percent }
\newlabel{supplement-best-practice}{{9}{39}{Software engineering best practice}{section.9}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {9.a}Requirements}{39}{subsection.9.1}\protected@file@percent }
\citation{differential-equations}
\citation{cbc}
\citation{spark}
\citation{misra}
\citation{hoare}
\citation{hoare}
\citation{numerals}
\citation{heedless}
\@writefile{toc}{\contentsline {subsection}{\numberline {9.b}Formal methods}{40}{subsection.9.2}\protected@file@percent }
\@writefile{toc}{\contentsline {subsection}{\numberline {9.c}Defensive programming}{40}{subsection.9.3}\protected@file@percent }
\citation{onlyPaperWithChecks}
\citation{tweet,plos}
\citation{heedless}
\citation{spark}
\citation{haskell}
\citation{basic-reproducibilty,reproducibility}
\citation{NVP}
\citation{avianFluModel}
\citation{ICmodel}
\citation{flu-model}
\citation{excel1,excel2}
\citation{npm}
\@writefile{toc}{\contentsline {subsection}{\numberline {9.d}Using inappropriate programming languages undermines reliability}{41}{subsection.9.4}\protected@file@percent }
\@writefile{toc}{\contentsline {subsection}{\numberline {9.e}Open source and version control}{41}{subsection.9.5}\protected@file@percent }
\citation{open-source}
\citation{sommerville,knight,NVP}
\citation{assurance-case}
\citation{LP}
\citation{relit}
\citation{hci1,hci2}
\@writefile{toc}{\contentsline {subsection}{\numberline {9.f}Rigorous testing}{42}{subsection.9.6}\protected@file@percent }
\@writefile{toc}{\contentsline {subsection}{\numberline {9.g}Good documentation and record keeping}{42}{subsection.9.7}\protected@file@percent }
\citation{flexibility}
\citation{iso}
\citation{178C}
\citation{fixit}
\@writefile{toc}{\contentsline {subsection}{\numberline {9.h}Usability}{43}{subsection.9.8}\protected@file@percent }
